Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-07-09 paper/lostanlenRelevanceScattering.tex
cite sainath 2015 interspeech
cite the Gammatone benchmark in lostanlen2017phd

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
Oscillations (red, blue) are the real and imaginary parts. The envelope (yellow) is the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment~\cite{Venkitaraman2014}, and are shown in Figure~\ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelopes of Gammatone wavelets are more biologically plausible than the symmetric, Gaussian envelopes of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking~\cite{Fastl2007}.
The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
Likewise, because critical bands are skewed towards higher frequencies, a masking tone has to be louder if it is above the stimulus in frequency rather than below.

It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds~\cite{Smith2006}.
This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.

>>>>

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
Oscillations (red, blue) are the real and imaginary parts. The envelope (yellow) is the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment~\cite{Venkitaraman2014}, and are shown in Figure~\ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelopes of Gammatone wavelets are more biologically plausible than the symmetric, Gaussian envelopes of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking~\cite{Fastl2007}.
The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
Likewise, because critical bands are skewed towards higher frequencies, a masking tone has to be louder if it is above the stimulus in frequency rather than below.
It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds~\cite{Smith2006,sainath2015interspeech}.
In addition, Gammatone wavelets have proven to outperform Morlet wavelets on a benchmark of supervised musical instrument classification from scattering coefficients \cite{lostanlen2017phd}.
This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.
